---
title:  "Process AK FRP data"
author: "JJWalker"
date: "2018-12-01"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---


####_process_ak_frp_data.Rmd_

This script inputs annual MxD14A1 FRP datasets produced in GEE and outputs an Rdata file for further plotting and analysis.  

_Input files_:  mxda1_gee_<yyyy>_plus_fire_info.csv, where <yyyy> = 2002-2016  

_Output file_:  ak_frp_area_data.RData

```{r, echo=FALSE}
# Prior to input, FRP files are processed as follows: 

#  * Formatted from GEE format to ArcMap-ready (_format_gee_file_for_import_into_arc.R_)
#  * Intersected with the BLM fire history file (_intersect_frp_data_with_fire_perimeters.py_) to combine 
#    * FRP data
#    * EVT data (in the coincident MODIS 1-km pixel)
#    * Fire history information
```


```{r, echo = FALSE, message=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
#### Set up ####
# ---------------------------------------------------------------- #

# Load libraries
library(lubridate) 
library(tidyverse)
library(dtplyr)
library(kableExtra) #kable tables
library(knitr) #kable

# Set paths
#path.in <- "D:/projects/ak_fire"
path.in = "/Users/jessicawalker/Documents/projects/ak_frp"
#path.r <- "D:/projects/ak_fire/data"
path.r <- file.path(path.in, "data")
rdata <- "ak_frp_data_pixel_area.RData"
#path.plots <- "D:/projects/ak_fire/output/plots/frp"
path.plots <- file.path(path.in, "output/plots")

source(file.path(path.in, "R", "ak_functions.R"))

```


```{r, echo = F}

# ---------------------------------------------------------------- #
#### Combine annual files ####
# ---------------------------------------------------------------- #

# Get the relevant filenames
  filenames.in <- list.files(file.path(path.in, "data/intersected"), pattern = "*plus_fire_info_bufferIn600m.csv")
  
# Read all files to a single file. Suppress warnings b/c of spurious "incomplete last line" error
  suppressWarnings(files.all <- do.call(rbind, lapply(file.path(path.in, "data/intersected", filenames.in), read.csv, row.names=NULL)))

```

```{r, echo = F}

# ---------------------------------------------------------------- #
#### Define columns ####
# ---------------------------------------------------------------- #

# Retain only specified columns
  colNames <- c("index",
  "class11", "class12", "class21", "class2197", "class22", "class23", "class24",     
  "class2600", "class2601", "class2602", "class2603", "class2604", "class2605", "class2606", "class2607", "class2608", "class2609", 
  "class2610", "class2611", "class2612",
  "class2631", "class2633", "class2634", "class2635", "class2636", "class2638", "class2639", 
  "class2640", "class2642", "class2643", "class2644", "class2645", "class2646", "class2648", "class2649", 
  "class2651", "class2652", 
  "class2671", "class2677", "class2678", "class2679", 
  "class2682", "class2683", "class2684", "class2685", "class2686", "class2687", "class2688", "class2689", 
  "class2690", "class2691", "class2692", "class2699", 
  "class2709", "class2718", "class2719", 
  "class2720",
  "class2730", 
  "class2740", "class2741", "class2742", "class2743", "class2744", "class2745", "class2746", "class2747",
  "class2751", "class2753", "class2756", "class2757", "class2758",     
  "class2761", "class2762", "class2763", "class2764", 
  "class2771", "class2772", "class2773", "class2774", "class2776", "class2777", 
  "class2781", "class2782", "class2783", "class2784", "class2785", "class2786",     
  "class2791", "class2792", "class2793", "class2794", 
  "class31", 
  "class81", "class82",
  "MaxFRP", "latitude", "longitude",
  "newdate", "FireYear", "FIREID", "DiscDate")

 df.orig <- files.all[, colNames]
 
```


```{r, echo = F}

# ---------------------------------------------------------------- #
#### Define EVT columns ####
# ---------------------------------------------------------------- #

# Define EVT classes and consolidated groups
  evt.classes <- data.frame(evt_number =  
                c("class11", "class12", "class21", "class2197","class22","class23", "class24",
                  "class2600", "class2601","class2602","class2603", "class2604","class2605","class2606","class260","class2608","class2609",
                  "class2610","class2611","class2612",
                  "class2631","class2633","class2634","class2635","class2636","class2638", "class2639",
                  "class2640", "class2642", "class2643", "class2644", "class2645", "class2646", "class2648", "class2649",
                  "class2651", "class2652",
                  "class2671","class2677","class2678", "class2679",
                  "class2682", "class2683", "class2684", "class2685","class2686", "class2687", "class2688", "class2689",
                  "class2690", "class2691", "class2692", "class2699",
                  "class2709", "class2718", "class2719", 
                  "class2720", 
                  "class2730",
                  "class2740","class2741", "class2742","class2743","class2744","class2745","class2746", "class2747",
                  "class2751","class2753","class2756", "class2757", "class2758",
                  "class2761", "class2762", "class2763","class2764",
                  "class2771", "class2772", "class2773", "class2774", "class2776","class2777",
                  "class2781", "class2782","class2783", "class2784", "class2785", "class2786", 
                  "class2791", "class2792", "class2793", "class2794","class31", 
                  "class81", "class82"),

  evt_group = c("Water", "Snow", "Developed", "Burned", "Developed", "Developed", "Developed",
                   "WhtSpruce", "WhtSpruce", "BlkSpruce", "WhtSpruce", "BlkSpruce", "BirchAspen", "Aspen", 
                   "BalsPopAsp", 
                   "Shrubland", "Shrubland", "Shrubland", "Grassland", "Grassland","Shrubland", 
                   "Grassland", 
                   "Shrubland", "Shrubland", "Shrubland", "Shrubland", "Shrubland", "Shrubland", 
                   "BirchAspen", "Shrubland", "Spruce", "Grassland", 
                   "Hemlock", "Hemlock", "Hemlock", 
                    "Grassland", "Shrubland", "Grassland",
                   "WhtSpruce", "WhtSpruce", "WhtSpruce", "Shrubland", 
                   "Tundra", "Tundra", "Tundra", "Tundra", "Tundra",
                   "Shrubland", "Shrubland", "Shrubland", 
                   "Tundra", "Tundra", 
                   "Grassland", "Grassland", 
                   "Shrubland", "Shrubland", "Shrubland", "Peatland", "Marsh", "Tidal",
                   "Tidal", "Wetland", "Wetland", "Wetland", "Wetland", "Grassland", "Wetland",
                   "BlkSpruce", "Shrubland", "Shrubland", "Shrubland", "Floodplain", "Floodplain", "Floodplain",
                   "Floodplain", "Peatland", "Peatland","Peatland", "Peatland", "Shrubland", "Swamp",
                   "Tundra", "Tundra","Tundra","Tundra","Tundra","Tundra","Barren", "Barren", "Barren",
                   "Barren", "Barren", "Agriculture", "Agriculture"))

```


```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### File clean up and formatting ####
# ---------------------------------------------------------------- #

# Point ID ('index') is unique to each point, though the same ID can occur in multiple rows since
# a single point may have burned in multiple fires. The input format is unwieldy; replace with 
# a simpler sequence of sequential numbers.
  levels(df.orig$index) <- c(seq(1:length(levels(df.orig$index))))
  
# Rename columns  
  oldnames = c("index","MaxFRP", "FireYear", "FIREID", "latitude", "longitude")
  newnames = c("ptid", "frp", "fire_year", "fireid", "lat", "lon")
  df.orig <- df.orig %>% rename_at(vars(oldnames), ~ newnames)

# Incoming MODIS frp is scaled by 10; rescale
  df.orig$frp <- 0.1 * df.orig$frp
  
# Remove NA-only columns
  df.orig <- Filter(function(y) !all(is.na(y)), df.orig)

# Remove rows in which frp == NA
  df.orig <- df.orig[!is.na(df.orig$frp), ]

# Format date columns and remove old ones
  df.orig$modis_date <- as.Date(df.orig$newdate, format = "%m/%d/%Y")  #"%Y-%m-%d")
  df.orig$fire_date <- as.Date(df.orig$DiscDate, format = "%m/%d/%Y")
  df.orig$modis_year <- year(df.orig$modis_date)
  
# Ensure point id is a factor
  df.orig$ptid <- factor(df.orig$ptid)

# Remove unneeded columns
  df.orig <- df.orig %>% select (-c(DiscDate, newdate))
  
# Remove all rows in which the modis date preceeds the fire date
   df.orig <- df.orig[which(df.orig$modis_date > df.orig$fire_date), ]
```
###Assemble new dataframe of fires

```{r, echo=FALSE, warning=F}

# ---------------------------------------------------------------- #
#### Set up the new fire df ####
# ---------------------------------------------------------------- #

# We need a row for each unique fire_year and modis_year.
# fire_years are derived from the BLM fire history database; modis_years are from FRP data.
# Both are necessary since some modis_years do not have a corresponding BLM fire_year.
# Each will have an FRP value for each MODIS year; pre-MODIS years get frp = NA.

# Set up df to hold entries for all points
  df.holder <- setNames(data.frame(matrix(ncol = 6, nrow = 0)),  c('ptid', 'lat', 'long', 'fire_year', 'frp', 'fireid'))

# Remove classes temporarily
  df.noclasses <- df.orig %>% select(-starts_with('class'))
  
# Create 'classes' df with 1 entry for each ptid
  df.classes <- df.orig %>% 
                select(ptid, starts_with('class')) %>% 
                distinct(ptid, .keep_all = T)

```
Example of a point with multiple associated fires (without class info):

`r df.noclasses %>%
          filter(ptid == 5767) %>% 
          arrange(fire_year) %>% 
          kable() %>%
          kable_styling()`

```{r, echo=FALSE}
# ---------------------------------------------------------------- #
#### Function to create new fire df ####
# ---------------------------------------------------------------- #

# For each ptid:
# - get unique years from combined set of fire years and modis years
#   (both are necessary b/c some modis years are not associated with a BLM fire) 
# - construct a new dataframe with invariant info (lat, long, evt)
# - replicate for the number of years 

create_new_fire_df = function(f) {

  # Trim factor levels
  f$ptid <- factor(f$ptid)

  # Create list of all unique fire and MODIS years. 
  years <- unique(c(f$fire_year, f$modis_year))
  
  # Set up new df with invariant point info for given # of years
  f.new <- f[1, c("ptid", "lat", "lon")] %>% slice(rep(1:n(), each = length(years)))

  # Add fire years
  f.new <- cbind(f.new, fire_year = years)

  # Get the frp corresponding to the modis year. Years preceeding modis era get NA
  # 'distinct' is to catch duplicates of modis year/frp, since the same frp is recorded for 
  # multiple fire years that preceed the modis era
  merge1 <- left_join(f.new, distinct(dplyr::select(f, modis_year, frp)), by = c("fire_year" = "modis_year"))
     
  # Join to get fire_id, etc.
  merge2 <- merge1 %>% left_join(distinct(dplyr::select(f, fireid, fire_year), fire_year, .keep_all = T), by = c("fire_year"))
  
  # Consolidate wth holder dataframe
   df.holder <- rbind(df.holder, merge2)
   
   return(df.holder) 
  }
```  

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

# ---------------------------------------------------------------- #
#### Create new fire dataframe ####
# ---------------------------------------------------------------- #

# ** THIS TAKES A WHILE **

# apply fcn to each ptid faction
lst.fires <- df.noclasses %>% 
            group_by(ptid) %>%
            do(create_new_fire_df(.))

# Convert from df of lists to just a df
df.fires.noevt <- data.frame(lst.fires)
df.fires.noevt <- df.fires.noevt[order(df.fires.noevt$ptid, df.fires.noevt$fire_year), ]

```

  
```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Function - Calculate burn number ####
# ---------------------------------------------------------------- #

# Count the number of times each point appears: 1 = first fire, 2 = 2nd, etc.
  get_fire_order = function(df) {
    df <- df %>% 
      group_by(ptid) %>% 
      arrange(fire_year) %>%  # put in ascending order
      mutate(burn_num = seq_along(fire_year)) # assign a sequence number
    df <- data.frame(df)
    df$reburn <- 0 
    df[which(as.numeric(df$burn_num) > 1), ]$reburn <- 1 # reburn: no = 0, yes = 1
    df$year_int <- ave(df$fire_year, factor(df$ptid), FUN=function(t) c(0, diff(t))) # get years between fires
  return(df)
  }

```

```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Get burn number for each point ####
# ---------------------------------------------------------------- #

df.fires.noevt <- get_fire_order(df.fires.noevt)
```

After processing, the same point id has this format:

`r df.fires.noevt %>%
  filter(ptid == 5767) %>% 
  kable() %>%
  kable_styling()`
  
Notes:

  * FRP = NA if the fire was prior to the MODIS era (documented only through the BLM history database)  
  * FRP pixels without a corresponding BLM database fire have BLM fire ID = NA, FRP = value 
  * BLM database fires in the MODIS era without a registered FRP value are removed

###EVT classes

After EVT classes are added back to the dataframe:

1. Determine EVT class with the greatest proportion in each 1-km MODIS pixel
2. Calculate area of that EVT class in each pixel
3. Remove pixels in which none of the classes had a majority

```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Merge fire df with EVT classes ####
# ---------------------------------------------------------------- #

# Necessary housekeeping
df.fires.noevt$ptid <- as.factor(df.fires.noevt$ptid)
df.classes$ptid <- as.numeric(df.classes$ptid)

# Merge
df.fires.evt <- merge(df.fires.noevt, df.classes, by = c("ptid"), all.x = T)

```


```{r, echo = FALSE, warning = F}

# ---------------------------------------------------------------- #
#### Process EVT classes ####
# ---------------------------------------------------------------- #
  
  # Record the EVT class that had the greatest area in each 1-km MODIS pixel
  df.fires.evt$max_evt_cat <- apply(df.fires.evt, 1, function(y) names(y[grepl("class", names(y))])[which.max(y[grepl("class", names(y))])[1]])
  
  # ...record the area that class represented
  df.fires.evt$max_evt_area <- apply(df.fires.evt[grepl("class", names(df.fires.evt))], 1, max, na.rm = TRUE)
  
  # ...calculate the proportion of pixels that class represented
  #df.fires.evt$max_evt_prop <- df.fires.evt$max_evt_pix/max_landsat_pixels * 100

  # Remove rows in which max_class == NA; this means none of the classes had a majority
  df.fires.evt <- df.fires.evt[!is.na(df.fires.evt$max_evt_cat), ]

  # Convert to factor
  df.fires.evt$max_evt_cat <- as.factor(df.fires.evt$max_evt_cat)

  # Merge with file of EVT classes to associate a name with each class
  df.fires.evt <- merge(df.fires.evt, evt.classes, by.x = "max_evt_cat", by.y = "evt_number")
  
  # Get rid of all individual classes. dplyr rocks!
  df.fires.evt <- df.fires.evt %>% select(-starts_with('class'))
  
  # Convert burn_num to factor
  df.fires.evt$burn_num <- as.factor(df.fires.evt$burn_num)
  
  # Rearrange 
  df.fires.evt <- df.fires.evt %>% select(ptid:evt_group, max_evt_cat)

```
`r df.fires.evt %>%
  filter(ptid == 5767) %>% 
  arrange(fire_year) %>% 
  kable() %>%
  kable_styling()`
  
<br>


###Summary stats
  
```{r, echo = F}

# ---------------------------------------------------------------- #
#### Calculate summary stats ####
# ---------------------------------------------------------------- #

# FRP by burn count (data)
x.frp.summary <- df.fires.evt %>%
                 group_by(burn_num) %>%
                 summarize(mean = mean(frp, na.rm = T),
                           med = median(frp, na.rm = T),
                           n = length(burn_num))

# > x.frp.summary
# # A tibble: 6 x 4
#   burn_num  mean   med     n
#   <fct>    <dbl> <dbl> <int>
# 1 1        395.  180.  43684
# 2 2        259.  125.  11391
# 3 3        273.   97.6  1169
# 4 4        147.   79.6   164
# 5 5         89.0  79.6    59
# 6 6         77.2 100.      8
```
####Overall FRP summary
`r x.frp.summary %>%
      kable(digits = 1) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)`

```{r, echo=FALSE}

# Get # of EVT classes with majorities
x.maxevt.gp <- df.fires.evt %>% 
                  group_by(evt_group, burn_num) %>%
                  summarize (mean = mean(frp, na.rm = T), 
                            med = median(frp, na.rm = T), 
                            n = length(burn_num))

# x.maxevt.gp <- data.frame(x.maxevt.gp)
# > x.maxevt.gp
#      evt_group burn_num      mean    med     n
# 1  Agriculture        1       NaN     NA    13
# 2  Agriculture        2  79.04167  28.15    13
# 3  Agriculture        3  68.20000  68.20     1
# 4        Aspen        1 224.40000 224.40     6
# 5        Aspen        2 124.22000  22.10     5
# 6       Barren        1 262.95556 142.35    21
# 7       Barren        2  48.83333  53.30     3
# 8   BirchAspen        1 353.46939 191.20  5165
# 9   BirchAspen        2 266.03183 134.70  2278
# 10  BirchAspen        3 357.14643 111.75   308
# 11  BirchAspen        4 357.15610 147.90    43
# 12  BirchAspen        5  60.28571  67.80     7
# 13  BirchAspen        6  89.27500 100.50     4
# 14   BlkSpruce        1 517.00040 236.40  7720
# 15   BlkSpruce        2 358.18542 179.90  1311
# 16   BlkSpruce        3 353.20928 254.60    97
# 17      Burned        1 120.76000 113.40    37
# 18      Burned        2 182.91111 204.60    27
# 19      Burned        3  95.96667  81.30     9
# 20  Floodplain        1 230.75295 121.00  2422
# 21  Floodplain        2 210.15423 101.90   747
# 22  Floodplain        3 134.63704  49.10    27
# 23   Grassland        1 531.41340 271.00   343
# 24   Grassland        2 210.39545 135.90    53
# 25   Grassland        3 410.67778 528.40     9
# 26       Marsh        1  74.60000  74.60     1
# 27    Peatland        1 162.73345  97.80   465
# 28    Peatland        2 145.04383  66.15   169
# 29    Peatland        3 125.45714  15.00     7
# 30   Shrubland        1 327.16200 155.40  3139
# 31   Shrubland        2 187.77614 102.25  1106
# 32   Shrubland        3 157.30638  78.60   152
# 33   Shrubland        4  51.99545  57.55    22
# 34   Shrubland        5  85.07500  77.40     4
# 35        Snow        1 155.43333 159.00    15
# 36        Snow        2 518.83333 273.70     9
# 37       Swamp        1  46.71667  31.90    16
# 38       Swamp        2  43.47143  45.10    10
# 39       Swamp        3  70.40000  70.40     3
# 40      Tundra        1 269.47828 129.00  4269
# 41      Tundra        2 210.48984 112.20  1964
# 42      Tundra        3  92.84247  56.70   312
# 43      Tundra        4  73.30278  79.60    81
# 44      Tundra        5  93.81053  79.60    39
# 45      Tundra        6  79.96667 109.60     3
# 46       Water        1 143.24211  80.80   112
# 47       Water        2 187.54857 124.00    36
# 48       Water        3  69.25000  69.25     2
# 49     Wetland        1 339.95167 157.50 10165
# 50     Wetland        2 252.14070 116.80  2048
# 51     Wetland        3 134.20115  70.40    99
# 52     Wetland        4  74.96154  68.70    16
# 53     Wetland        5  92.52222 109.60     9
# 54     Wetland        6  20.60000  20.60     1
# 55   WhtSpruce        1 461.96088 219.25  9775
# 56   WhtSpruce        2 312.68073 152.40  1612
# 57   WhtSpruce        3 624.28380 387.70   143
# 58   WhtSpruce        4  32.05000  32.05     2


# Keep only those levels that have at least 10 points
x.maxevt.gp.top <- x.maxevt.gp[x.maxevt.gp$n > 10, ]

# Keep only those classes that have at least 2 burn levels
x.maxevt.gp.top <- x.maxevt.gp.top[!(as.numeric(x.maxevt.gp.top$evt_group) %in% which(table(x.maxevt.gp.top$evt_group) < 2)), ]

```

####FRP summary by primary EVT class
`r x.maxevt.gp.top %>%
      kable(digits = 1) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)`

####Save data

RData file is saved to `r file.path(path.r, rdata)`
```{r, echo=FALSE}

# Save data and environment settings   
print(paste0("R data file saved to ", file.path(path.r, rdata)))
save.image(file = file.path(path.r, rdata))
