---
title: ## "FRP burn data plots"
author: "JWalker"
date: "2018-07-27"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---
```{r, echo=FALSE, warning=FALSE, include=FALSE}
########################################################################### #
#
# plot_ak_frp_data.R 
#
# Objective:  Plot processed MxD14A1 FRP data.
#
# Input:     .RData file produced by process_ak_frp_data.R
#
# Output:     Box-plots of FRP by burn frequency (1x, 2x, 3x burns)
#             
#   
#   JWalker 17 April 2018
#  
########################################################################### #

# ---- Set environment ----

# Remove existing data and values
rm(list=ls())

# Load libraries
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr) # for formatting tables
library(car) #for unbalanced ANOVA test

# Set folder paths
path.in <- "D:/projects/ak_fire" 
path.plots <- "D:/projects/ak_fire/output/plots/frp"

# Load functions
source(file.path(path.in, "R", "ak_functions.R"))

# Load data saved from 'process_ak_frp_data.R'
rdata <- "ak_frp_data.RData"
# https://drive.google.com/open?id=1w1O-Rife6raqvYMBWjTFrCT8IEm_PZM7

data_url = "1w1O-Rife6raqvYMBWjTFrCT8IEm_PZM7"
df <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", data_url))
```

```{r, echo=FALSE}
load(file = file.path(path.in, "data", rdata))

# Set consistent plot parameters
dodge = position_dodge(0.16)
theme_set(theme_bw())
plot_opts <- theme(panel.grid.minor.x = element_blank(),
                   panel.grid.major.x = element_blank(), # hide gridlines
                   legend.key = element_blank(),  # remove boxes around legend items
                   plot.title = element_text(hjust = 0.5)) # center title
```

### Plots related to the analysis of FRP data in Alaskan fires
<br>
Summary of FRP values according to primary EVT group in a given pixel and burn number. Here only EVT classes with > 10 observations and > 1 burn level are considered.

```{r, echo=FALSE, include=FALSE}
x.maxevt.gp <- x %>% 
                  group_by(evt_group, burn_num) %>%
                  summarize (mean_FRP = mean(MaxFRP), 
                            median_FRP = median(MaxFRP), 
                            sd_FRP = sd(MaxFRP, na.rm = TRUE),
                            n = length(burn_num))

# Keep only those levels that have at least 10 points
x.maxevt.gp.top <- x.maxevt.gp[x.maxevt.gp$n > 10, ]

# Keep only those classes that have at least 2 burn levels
x.maxevt.gp.top <- x.maxevt.gp.top[!(as.numeric(x.maxevt.gp.top$evt_group) %in%     which(table(x.maxevt.gp.top$evt_group) < 2)), ]

# Make subset of those classes
x.evtsub <- subset(x, (evt_group %in% x.maxevt.gp.top$evt_group))
x.evtsub$evt_group <- droplevels(x.evtsub$evt_group)

# Merge tables to associate an n with each value
x.evtsub.top <- merge(x.evtsub, x.maxevt.gp.top, by = c("burn_num", "evt_group"))

# Get rid of all rows where n < 10
x.evtsub.top <- x.evtsub.top[which(x.evtsub.top$n > 9),]

# Show pretty table
kable(x.maxevt.gp.top, caption = "FRP summary")

```


Given the non-normality of FRP values, log(FRP) transformations are appropriate.

```{r, echo=FALSE, include=FALSE}

# ------- BOXPLOT by burn number, non-log plot

title <- "FRP by burn number, all years, combined land cover"
plot.name <- "AK_FRP by burn number_all years.png"

# Get means
means <- aggregate(MaxFRP ~ burn_num, x, mean)

# Plot boxplot with mean values 
p <- ggplot(x, aes(as.factor(burn_num), MaxFRP)) + 
            geom_boxplot(aes(fill = as.factor(burn_num)), show.legend = F) +
            geom_point(data=x.frp.summary, aes(as.factor(burn_num), mean, shape = "mean"), size = 7) +
            scale_shape_manual("", values = 18) +
            geom_text(data = means, aes(label = round(MaxFRP, 2), y = MaxFRP + 16), 
                      hjust = -0.4, vjust = -0.3) +
            coord_cartesian(ylim = c(0,500), expand = TRUE) +
            labs(x = "Burn number", y = "FRP") +
           ggtitle(title)
 p + plot_opts

# ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)

```

```{r, echo = FALSE, eval=TRUE}

# ------- BOXPLOT by burn number, log plot

title <- "Log FRP by burn number, all years, combined land cover"
plot.name <- "AK_FRP by burn number_all years_log.png"


# Plot boxplot with mean values 
p <- ggplot(x.evtsub.top, aes(as.factor(burn_num), log10(MaxFRP))) + 
  geom_boxplot(aes(fill = as.factor(burn_num)), show.legend = F) +
  geom_point(data=means, aes(as.factor(burn_num), log10(MaxFRP), shape = "mean"), size = 7) +
  scale_shape_manual("", values = 18) +
  geom_text(data = means, aes(label = round(log10(MaxFRP), 2), y = log10(MaxFRP) + 0.1), 
            hjust = -0.4, vjust = 0.1) +
  labs(x = "Burn number", y = "log FRP") +
  ggtitle(title)


p + plot_opts

#ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)


```
Below is the same data broken out by land cover, where "land cover" is assigned to whichever 2001 EVT class covered the greatest proportion of each 1-km MODIS pixel. 
```{r, echo=FALSE, eval=TRUE}

#--------- BOXPLOT by max vegetation classes

title <- "Log FRP by maximum vegetation classburn number, all years"
plot.name <- "AK_FRP by burn number_all years_log.png"
p <- ggplot(x.evtsub.top, aes(burn_num, log10(MaxFRP))) + 
      geom_boxplot(aes(fill=burn_num)) + 
      labs(x = "Burn number", y = "log FRP") +
      facet_wrap( ~ evt_group, ncol = 3) +
      theme(legend.position = "none")
#      coord_cartesian(ylim = c(0, 800), expand = TRUE)
p + plot_opts

```

Alternate view:
```{r, echo=FALSE}

interaction.plot(x.evtsub.top$burn_num, x.evtsub.top$evt_group, log10(x.evtsub.top$MaxFRP), type = "b", col=c(1:3), xlab = "Burn number", ylab = "Mean log FRP", main = "Interaction plot",trace.label = "EVT group")


```
Black Spruce, Burned, Tundra, and Wetland appear to have decreasing FRP relationships with burn number. White Spruce has an increasing trend. Throwing everything into a 2-way ANOVA:

```{r, echo=FALSE}
# Calculate 2-way ANOVA; evt group and burn num
x.anova <- aov(formula = MaxFRP ~ evt_group * burn_num, data = x.evtsub.top)
summary(x.anova)
```

```{r,, echo=FALSE}

TukeyHSD(x.anova, which = "burn_num")
```
Despite what appears to be more dramatic changes in FRP after burn #2, the only identified significance is between burns #1 and #2. This is consistent with the individual anova tests of the influence of burn number; Shrubland and White Spruce are the only classes that show significant differences between burns--specifically, burns #1 and #2. 

<br>
Shrubland
```{r, echo=FALSE}

test <- aov(formula = MaxFRP ~ burn_num, data = subset(x.evtsub.top, evt_group == "Shrubland"))
summary(test)

```
```{r, echo=FALSE}

TukeyHSD(test, which = "burn_num")

```
<br>

White Spruce
```{r, , echo=FALSE}
test <- aov(formula = MaxFRP ~ burn_num, data = subset(x.evtsub.top, evt_group == "WhtSpruce"))
summary(test)

```
```{r, , echo=FALSE}

TukeyHSD(test, which = "burn_num")

```




### Influence of years since prior burn on FRP

Plot of all combined classes:
<br>

```{r, echo=FALSE}

x.yr <- x

# ------ PLOT by years since prior burn
# Drop factors for which there are fewer than 10 points
  keep <- levels(as.factor(x$year_int))[table(x$year_int) > 10]

  title <- "FRP by years since prior burn"
  p <- ggplot(subset(x.evtsub, year_int > 0), aes(year_int, log10(MaxFRP), group = year_int, fill = year_int)) + 
       geom_boxplot() + 
       labs(x = "Years since prior burn", y = "log FRP") +
       ggtitle(title) +
       scale_fill_continuous(name = "Years since fire")
  p + plot_opts

  #ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)

```

Fitting a line to the means shows a slight upward trend in overall FRP as time between burns increases, but low R<sup>2</sup>.

```{r, echo=FALSE}
  
# -------- PLOT by mean fire interval
  x.summary <- x.evtsub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  title <- paste0("Mean FRP by # years since prior burn")

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "Mean log FRP") +
       annotate("text", x = 2, y = 0.98*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

#ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)
```
By individual EVT class. Only significant (p < 0.05) results are shown.

#### **Birch/Aspen**
```{r, echo=FALSE, warning=FALSE}
  
title <- "Mean FRP by # of years since prior burn"

# -------- PLOT by mean fire interval
  x.sub <- subset(x.evtsub.top, evt_group == "BirchAspen" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)
  
  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "Mean log FRP") +
       annotate("text", x = 2, y = 0.98*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  
  
  p + plot_opts


```

 
```{r, echo=FALSE, include=FALSE}

# BLACK SPRUCE
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "BlkSpruce" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.9*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

```
#### **Burned**
```{r, echo=FALSE}
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x.evtsub.top, evt_group == "Burned" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.98*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts
```


```{r, echo=FALSE,  include=FALSE}

# FLOODPLAIN
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Floodplain" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.9*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts


```

 
```{r, echo=FALSE, include=FALSE}

# GRASSLAND
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Grassland" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.9*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts


```


```{r, echo=FALSE,  include=FALSE}

# PEATLAND
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Peatland" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.9*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

```


```{r, echo=FALSE, include=FALSE}

# SHRUBLAND  

# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Shrubland" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.99*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

```


```{r, echo=FALSE,include=FALSE}

# SWAMP

# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Swamp" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 30, y = 0.9*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

```
#### **Tundra**
```{r, echo=FALSE}

# TUNDRA
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Tundra" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.98*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts


```
#### **Wetland**
```{r, echo=FALSE}

# WETLAND
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "Wetland" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

  p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 2, y = 0.98*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

```

#### **White Spruce**
```{r, echo=FALSE, warning=FALSE}

# WHITE SPRUCE
  
# -------- PLOT by mean fire interval

  x.sub <- subset(x, evt_group == "WhtSpruce" & year_int != 0)
  x.summary <- x.sub %>% group_by(year_int) %>% summarize(mean = mean(log10(MaxFRP)))

  y.max <- max(x.summary$mean)

    p <- ggplot(x.summary, aes(year_int, mean)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 20, y = 0.96*y.max, label = lm_eqn(x.summary, "mean", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts


```

<br>
<br>

```{r, echo=FALSE, include=FALSE}


# -------- HISTOGRAM of distributions

p <- ggplot(x, aes(x=log10(MaxFRP), fill = burn_num)) + 
      geom_histogram(alpha = 0.5, position = "identity",
      binwidth = 0.1) +
      xlab("log FRP")

p + plot_opts

```






```{r, eval=FALSE, include=FALSE}

# Calculate semivariograms
x.copy <- x.yr

# convert degrees to km distances
x.copy$lon_km <- 110. * (x.copy$longitude - (-125))*cos(x.copy$latitude/(360/(2*pi)))
x.copy$lat_km <- 110. * (x.copy$latitude-65)
coordinates(x.copy) = ~lon_km+lat_km


# ------- PLOT points in x,y km space  
plot.name <- paste0("AK_FRP bubble plot ", yr, ".png")
png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(bubble(x.copy, zcol = 'MaxFRP', fill = T, do.sqrt = F, maxsize = 1.8))
dev.off()

# ------- PLOT variogram
v = variogram(MaxFRP~1, x.copy)
level <- c(0.5, 0.95)

plot.name <- paste0("AK_FRP variogram ", yr, ".png")
png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(plot(v, fraction = 0.65, level=level))
dev.off()


# # ------PLOT directional variogram
#   v.fit = fit.variogram(v, vgm("Sph"))
#   x.copy.var.det <- variogram(MaxFRP~lat_km+lon_km, data = x.copy)
#   
#   plot.name <- paste0("AK_FRP directional variogram ", yr, ".png")
#   p <- plot(x.copy.var.det)
#   p

plot.name <- paste0("AK_FRP directional variogram ", yr, ".png")

x.copy.var.dir <- variogram(MaxFRP~1, data = x.copy, alpha = c(0, 45, 90, 135))
#png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(plot(x.copy.var.dir))
dev.off()

}
```
