---
title:  "Plot AK FRP data"
author: "JJWalker"
date: "2016-08-01"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---
```{r, echo = FALSE}
########################################################################### #
#
# plot_ak_frp_data.Rmd
#
# Objective:  Plot processed MxD14A1 FRP data.
#
# Input:     .RData file produced by process_ak_frp_data.R
#
# Output:     FRP plots
#   
#   JWalker 17 April 2018
#  
########################################################################### #


# ---- Set environment ----

# Remove existing data and values
rm(list=ls())

# Load libraries
library(ggplot2)
library(lubridate)
library(kableExtra)
library(knitr)

# Set folder paths
path.in <- "/Users/jessicawalker/Documents/projects/ak_frp"
path.plots <- file.path(path.in, "/output/plots")  #"D:/projects/ak_fire/output/plots/frp"

# Load functions
source(file.path(path.in, "R", "ak_functions.R"))
```
####Data import

```{r, echo = F}
# Load data saved from 'process_ak_frp_data.R'
rdata <- "ak_frp_data_pixel_area.RData"
load(file = file.path(path.in, "data", rdata))

```

<br>
Data table is saved in this format:
`r df.fires.evt %>%
  arrange(fire_year) %>% 
  filter(ptid == 5767) %>% 
  kable() %>%
  kable_styling()`
```{r, echo = FALSE}
# ------------------------------------ -
# #### Set plot parameters ####
# ------------------------------------

# Set consistent plot parameters
dodge = position_dodge(0.16)
theme_set(theme_bw())
plot_opts <- theme(panel.grid.minor.x = element_blank(),
                   panel.grid.major.x = element_blank(), # hide gridlines
                   legend.key = element_blank(),  # remove boxes around legend items
                   plot.title = element_text(hjust = 0.5)) # center title

```

```{r, echo = FALSE, warning=FALSE}
# ---- All data, all years -----
# Only the highest burn count is retained to avoid overcounting contributions

# ------- BOXPLOT by burn number
title <- "FRP by burn number, all years, combined land cover"
plot.name <- "AK_FRP by burn number_all years.png"

# Get means
means <- aggregate(frp ~ burn_num, df.fires.evt, mean)

# Plot boxplot with mean values 
p <- ggplot(df.fires.evt, aes(as.factor(burn_num), frp)) + 
            geom_boxplot(aes(fill = as.factor(burn_num)), show.legend = F) +
            geom_point(data=x.frp.summary, aes(as.factor(burn_num), mean, shape = "mean"), size = 7) +
            scale_shape_manual("", values = 18) +
            geom_text(data = means, aes(label = round(frp, 2), y = frp + 16), hjust = -0.2) +
            coord_cartesian(ylim = c(0,500), expand = TRUE) +
            labs(x = "Burn number", y = "FRP") +
            ggtitle(title)
p + plot_opts

#ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)

```


```{r, echo= FALSE, warning=FALSE}
# ------- BOXPLOT by burn number, log plot

title <- "Log FRP by burn number, all years"
plot.name <- "AK_FRP by burn number_all years_log.png"


# Plot boxplot with mean values 
p <- ggplot(df.fires.evt, aes(as.factor(burn_num), log10(frp))) + 
  geom_boxplot(aes(fill = as.factor(burn_num)), show.legend = F) +
  geom_point(data=means, aes(as.factor(burn_num), log10(frp), shape = "mean"), size = 7) +
  scale_shape_manual("", values = 18) +
  geom_text(data = means, aes(label = round(log10(frp), 2), y = log10(frp) + 0.1), 
            hjust = -0.4, vjust = 0.1) +
  labs(x = "Burn number", y = "log FRP") +
  ggtitle(title)
p + plot_opts

#ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)

```


```{r, echo=FALSE, warning=FALSE}
#--------- BOXPLOT by max vegetation classes

df.sub <- subset(df.fires.evt, (evt_group %in% x.maxevt.gp.top$evt_group))

p <- ggplot(df.sub, aes(burn_num, log10(frp))) + 
      geom_boxplot() + 
      facet_wrap( ~ evt_group, ncol = 2) #+ 
#      coord_cartesian(ylim = c(0, 800), expand = TRUE)
p + plot_opts

```
```{r, echo=FALSE, warning=FALSE}
#--------- BOXPLOT by vegetation groups of interest

goi <- c('Aspen', 'BirchAspen', 'BlkSpruce', 'Grassland', 'Shrubland', 'Tundra', 'WhtSpruce')

df.sub <- subset(df.fires.evt, (evt_group %in% goi))

p <- ggplot(df.sub, aes(burn_num, log10(frp))) + 
      geom_boxplot() + 
      facet_wrap( ~ evt_group, ncol = 2) #+ 
#      coord_cartesian(ylim = c(0, 800), expand = TRUE)
p + plot_opts

```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# -------- HISTOGRAM of distributions
p <- ggplot(df.fires.evt, aes(x=log10(frp), fill = burn_num)) + 
      geom_histogram(alpha = 0.5, position = "identity") 

p + plot_opts
```




```{r, echo=FALSE}

df.sub <- subset(df.fires.evt, evt_group %in% goi)


df.ints <- 
   df.sub %>% filter(reburn == 1) %>% 
   mutate(ints = cut(year_int, c(seq(0, 70, 10)))) %>% 
   group_by(evt_group, ints) %>% 
   summarise(mean.frp = mean(frp, na.rm = T), 
             med.frp = median(frp, na.rm = T),
             n = length(frp)) %>% 
  filter(n > 10)

ggplot(df.ints, aes(ints, med.frp, group = evt_group, color = evt_group)) + geom_line() + geom_point()



```



```{r, echo=FALSE}
# For individual years. This mostly groups data geographically as well.

for (yr in seq(2002, 2016)) {
  
  x.yr <- subset(df.fires.evt, fire_year == yr)
  
# ------- PLOT by burn number
  title <- paste0("FRP by burn number, ", yr)
  plot.name <- paste0("AK_FRP by burn number_", yr, ".png")

  p <- ggplot(x.yr, aes(as.factor(burn_num), log10(frp))) + 
    geom_boxplot() + 
    #ylim(0, 1000) +
    labs(x = "Burn number", y = "log FRP") +
    ggtitle(title)
  p + plot_opts
#  ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)


# ------ PLOT by years since prior burn
# Drop factors for which there are fewer than 10 points
  keep <- levels(as.factor(x.yr$year_int))[table(x.yr$year_int) > 10]
  x.sub.int <- x.yr[x.yr$year_int %in% keep, ]

  title <- paste0("FRP by years since prior burn, ", yr)
  plot.name <- paste0("AK_FRP by years since prior burn_", yr, ".png")

  p <- ggplot(x.sub.int, aes(as.factor(year_int), log10(frp))) + 
       geom_boxplot() + 
       #ylim(0, 1000) + 
       labs(x = "Years since prior burn", y = "log FRP") +
       ggtitle(title)
  p + plot_opts

  ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)

# ------- PLOT by year of fire
  keep <- levels(as.factor(x.yr$FireYear))[table(x.yr$FireYear) > 10]
  x.sub.yr <- x.yr[x.yr$FireYear %in% keep, ]

  title <- paste0("FRP by year of fire, ", yr)
  plot.name <- paste0("AK_FRP by year of fire_", yr, ".png")

  p <- ggplot(x.sub.yr, aes(as.factor(FireYear), log10(MaxFRP))) + 
       geom_boxplot() + 
       #ylim(0, 1000) +
       labs(x = "Year of fire", y = "log FRP") +
       ggtitle(title)
  p + plot_opts

  ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)
  
}
```

```{r}
# -------- PLOT by mean fire interval
  x.pts <- x.sub.int %>% group_by(year_int) %>% summarize(avg = mean(frp))

x.pts <- subset(x, evt_group == 'WhtSpruce')
x.pts <-  x.pts %>% group_by(burn_num, year_int) %>% summarise(avg = mean(frp))

lm_eqn2 = function(df){
  m = lm(y ~ poly(x, 3), df)#3rd degree polynomial
  eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                   list(a = format(coef(m)[1], digits = 2),
                        b = format(coef(m)[2], digits = 2),
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq))
}

  #title <- paste0("Mean FRP by year interval, ", yr)
  #plot.name <- paste0("AK_FRP mean FRP by year ", yr, ".png")

  p <- ggplot(x.pts, aes(year_int, avg)) + 
       geom_point() + 
       geom_smooth(method = 'lm', se = FALSE, size = 0.4) +
       labs(x = "Years since prior burn", y = "FRP") +
       annotate("text", x = 5, y = 400, label = lm_eqn2(x.pts, "avg", "year_int"), 
           parse = TRUE, hjust = 0, vjust = 0, color = "#000000") +
       ggtitle(title)
  p + plot_opts

#ggsave(filename = plot.name, path = path.plots, width = 10, height = 7, units = c("in"), dpi = 600)
```

# Calculate semivariograms
x.copy <- x.yr

# convert degrees to km distances
x.copy$lon_km <- 110. * (x.copy$longitude - (-125))*cos(x.copy$latitude/(360/(2*pi)))
x.copy$lat_km <- 110. * (x.copy$latitude-65)
coordinates(x.copy) = ~lon_km+lat_km


# ------- PLOT points in x,y km space  
plot.name <- paste0("AK_FRP bubble plot ", yr, ".png")
png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(bubble(x.copy, zcol = 'MaxFRP', fill = T, do.sqrt = F, maxsize = 1.8))
dev.off()

# ------- PLOT variogram
v = variogram(MaxFRP~1, x.copy)
level <- c(0.5, 0.95)

plot.name <- paste0("AK_FRP variogram ", yr, ".png")
png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(plot(v, fraction = 0.65, level=level))
dev.off()


# # ------PLOT directional variogram
#   v.fit = fit.variogram(v, vgm("Sph"))
#   x.copy.var.det <- variogram(MaxFRP~lat_km+lon_km, data = x.copy)
#   
#   plot.name <- paste0("AK_FRP directional variogram ", yr, ".png")
#   p <- plot(x.copy.var.det)
#   p

plot.name <- paste0("AK_FRP directional variogram ", yr, ".png")

x.copy.var.dir <- variogram(MaxFRP~1, data = x.copy, alpha = c(0, 45, 90, 135))
png(file = file.path(path.plots, plot.name), width = 10, height = 7, units = c("in"), res = 600)
print(plot(x.copy.var.dir))
dev.off()

}


```{r, echo = FALSE}

# --------------------------------- 
# Stats
# ---------------------------------

# Calculate 2-way ANOVA; evt group and burn num
x.anova <- aov(formula = frp ~ evt_group * burn_num, data = df.fires.evt)
summary(x.anova)
TukeyHSD(x.anova, which = "burn_num")
TukeyHSD(x.anova, which = "evt_group")


# Calculate anova: FRP vs year_int, evt, and burn_num
x.anova <- aov(formula = frp ~ year_int * as.factor(evt_group) * burn_num, data = subset(df.fires.evt, reburn > 0))
summary(x.anova)

```