---
title: "R Notebook"
output: html_notebook
---

```{r, echo=FALSE}

########################################################################### #
#
# process_ak_frp_data.Rmd 
#
# Objective:  Process MxD14A1 FRP data in order to plot and analyze data.
#
# Input:     Files of MxD14A1 data originally produced in GEE
#            for fire complexes in Alaska and then processed via the following
#            steps: 
#             - formatted (format_gee_file_for_import_into_arc.R)
#             - intersected in Arc with the vector file of fire history 
#            (intersect_frp_data_with_fire_perimeters.py) to combine 
#            FRP data, EVT data (in the encompassing
#            MODIS 1-km pixel), and fire history data. The value in each EVT
#            class column is the area (m^2) of a given EVT class.
#
#   files:   mxda1_gee_<yyyy>_plus_fire_info.csv, where <yyyy> = 2002-2016
#
# Output:   ak_frp_area_data.RData.  Contains files processed through to 
#           statistical results.
#
#   JWalker 6 April 2018, revised substantially in Dec 2018
#  
########################################################################### #
```



```{r}

# Remove existing data and libraries from R environment
#rm(list=ls())

# Load libraries
library(lubridate) # dates
library(tidyverse)
library(dtplyr)

# Set paths
#path.in <- "D:/projects/ak_fire"
path.in = "/Users/jessicawalker/Documents/projects/ak_frp"
#path.r <- "D:/projects/ak_fire/data"
path.r <- file.path(path.in, "data")
rdata <- "ak_frp_data_pixel_area.RData"
#path.plots <- "D:/projects/ak_fire/output/plots/frp"
path.plots <- file.path(path.in, "output/plots")

source(file.path(path.in, "R", "ak_functions.R"))

```



```{r,echo=F}


# if the data are already saved in an Rdata file, skip this part
#if (!file.exists(file.path(path.r, rdata))){

  # Get the relevant filenames
#  filenames.in <- list.files(file.path(path.in, "data/tables"), pattern = "*plus_fire_info_bufferIn600m.csv")
  filenames.in <- list.files(file.path(path.in, "data/intersected"), pattern = "*plus_fire_info_bufferIn600m.csv")
  
  # Read all files to a single file. Suppress warnings b/c of spurious "incomplete last line" error
  suppressWarnings(x.all <- do.call(rbind, lapply(file.path(path.in, "data/intersected", filenames.in), read.csv,row.names=NULL)))
# this gets rid of the extra column
#  colnames(x.all) <- c(colnames(x.all)[-1],NULL)


```

```{r, echo=F}
  # Retain only specified columns
  colNames <- c("index",  #FID_firePe",
  "class11", "class12", "class21", "class2197", "class22", "class23", "class24",     
  "class2600", "class2601", "class2602", "class2603", "class2604", "class2605", "class2606", "class2607", "class2608", "class2609", 
  "class2610", "class2611", "class2612",
  "class2631", "class2633", "class2634", "class2635", "class2636", "class2638", "class2639", 
  "class2640", "class2642", "class2643", "class2644", "class2645", "class2646", "class2648", "class2649", 
  "class2651", "class2652", 
  "class2671", "class2677", "class2678", "class2679", 
  "class2682", "class2683", "class2684", "class2685", "class2686", "class2687", "class2688", "class2689", 
  "class2690", "class2691", "class2692", "class2699", 
  "class2709", "class2718", "class2719", 
  "class2720",
  "class2730", 
  "class2740", "class2741", "class2742", "class2743", "class2744", "class2745", "class2746", "class2747",
  "class2751", "class2753", "class2756", "class2757", "class2758",     
  "class2761", "class2762", "class2763", "class2764", 
  "class2771", "class2772", "class2773", "class2774", "class2776", "class2777", 
  "class2781", "class2782", "class2783", "class2784", "class2785", "class2786",     
  "class2791", "class2792", "class2793", "class2794", 
  "class31", 
  "class81", "class82",
  "MaxFRP", "latitude", "longitude",
  "newdate", "FireYear", "FIREID", "DiscDate")
```

```{r, echo=F}

 x <- x.all[, colNames]  

  # Set up EVT classes and consolidated groups
  evt_classes <- data.frame(evt_number =  
                c("class11", "class12", "class21", "class2197","class22","class23", "class24",
                  "class2600", "class2601","class2602","class2603", "class2604","class2605","class2606","class260","class2608","class2609",
                  "class2610","class2611","class2612",
                  "class2631","class2633","class2634","class2635","class2636","class2638", "class2639",
                  "class2640", "class2642", "class2643", "class2644", "class2645", "class2646", "class2648", "class2649",
                  "class2651", "class2652",
                  "class2671","class2677","class2678", "class2679",
                  "class2682", "class2683", "class2684", "class2685","class2686", "class2687", "class2688", "class2689",
                  "class2690", "class2691", "class2692", "class2699",
                  "class2709", "class2718", "class2719", 
                  "class2720", 
                  "class2730",
                  "class2740","class2741", "class2742","class2743","class2744","class2745","class2746", "class2747",
                  "class2751","class2753","class2756", "class2757", "class2758",
                  "class2761", "class2762", "class2763","class2764",
                  "class2771", "class2772", "class2773", "class2774", "class2776","class2777",
                  "class2781", "class2782","class2783", "class2784", "class2785", "class2786", 
                  "class2791", "class2792", "class2793", "class2794","class31", 
                  "class81", "class82"),

  evt_group = c("Water","Snow", "Developed", "Burned", "Developed", "Developed", "Developed",
                   "WhtSpruce", "WhtSpruce", "BlkSpruce", "WhtSpruce", "BlkSpruce", "BirchAspen", "Aspen", 
                   "BalsPopAsp", 
                   "Shrubland", "Shrubland", "Shrubland", "Grassland", "Grassland","Shrubland", 
                   "Grassland", 
                   "Shrubland", "Shrubland", "Shrubland", "Shrubland", "Shrubland", "Shrubland", 
                   "BirchAspen", "Shrubland", "Spruce", "Grassland", 
                   "Hemlock", "Hemlock", "Hemlock", 
                    "Grassland", "Shrubland", "Grassland",
                   "WhtSpruce", "WhtSpruce", "WhtSpruce", "Shrubland", 
                   "Tundra", "Tundra", "Tundra", "Tundra", "Tundra",
                   "Shrubland", "Shrubland", "Shrubland", 
                   "Tundra", "Tundra", 
                   "Grassland", "Grassland", 
                   "Shrubland", "Shrubland", "Shrubland", "Peatland", "Marsh", "Tidal",
                   "Tidal", "Wetland", "Wetland", "Wetland", "Wetland", "Grassland", "Wetland",
                   "BlkSpruce", "Shrubland", "Shrubland", "Shrubland", "Floodplain", "Floodplain", "Floodplain",
                   "Floodplain", "Peatland", "Peatland","Peatland", "Peatland", "Shrubland", "Swamp",
                   "Tundra", "Tundra","Tundra","Tundra","Tundra","Tundra","Barren", "Barren", "Barren",
                   "Barren", "Barren", "Agriculture", "Agriculture"))



```


```{r}

  # Index is unique to each point (though the same index can occur in multiple rows, since
  # points reference different fires) but cumbersome in its current form;
  # replace with sequence of sequential numbers. Each duplicate index will have a duplicate number.
  
  levels(x$index) <- c(seq(1:length(levels(x$index))))
  
  # Rename index
  colnames(x)[which(colnames(x) == "index")] <- "ptid"
  colnames(x)[which(colnames(x) == "MaxFRP")] <- "frp"
  colnames(x)[which(colnames(x) == "FireYear")] <- "fire_year"
  colnames(x)[which(colnames(x) == "FIREID")] <- "fireid"

  # Set FID as unique id for each row
  x$FID <- 1:nrow(x)

  # Reorder s.t. FID is first
  x <- x[, c(ncol(x), 1:(ncol(x)-1))]

  # Remove NA-only columns
  x <- Filter(function(y) !all(is.na(y)), x)

  # Remove rows in which frp == NA
  x <- x[!is.na(x$frp), ]

  # Format date columns and remove old ones
  x$modis_date <- as.Date(x$newdate, format = "%m/%d/%Y")  #"%Y-%m-%d")
  x$fire_date <- as.Date(x$DiscDate, format = "%m/%d/%Y")
  x$modis_year <- year(x$modis_date)
  
  # Ensure point id is a factor
  x$ptid <- factor(x$ptid)

  # Housekeeping
  x$DiscDate <- NULL
  x$newdate <- NULL

  # MODIS frp is scaled by 10
  x$frp <- 0.1 * x$frp
  
  x.backup <- x

```
  
```{r, echo  = FALSE}

  # Function to count the number of times each point appears: 1 = first fire, 2 = 2nd, etc.
  get_fire_order = function(df) {
  #  df <- df[order(df$fire_year, decreasing = FALSE),]
    df <- df %>% 
      arrange(fire_year) %>% 
      group_by(ptid) %>% 
      mutate(burn_num = seq_along(fire_year))
    df <- data.frame(df)
    df$reburn <- 0 
    df[which(as.numeric(df$burn_num) > 1), ]$reburn <- 1 #Reburn: no = 0, yes = 1
    df$year_int <- ave(df$fire_year, factor(df$ptid), FUN=function(t) c(0, diff(t))) #Get years between fires
  return(df)
  }
  
  

```

 
 
```{r}

  ###########################################################################
  # Here's where I should figure out how to branch off and then reconcile
  #
  # The attempt is to identify points for which there is an frp pixel in a given year but no recorded fire perimeter
  # Because these points are not associated with a named burn, they do not register as a fire. 
  # Get rid of all individual classes. Keep as separate file s.t. proportions
  # can be analyzed if desired
  
  
  # Get ptids for which more than 1 MODIS date exist
  #    t <- x.sub %>% group_by(ptid) %>% filter(length(unique(modis_date)) > 1)
   #   t <- data.frame(t)
  
  # Get ptids for which MODIS year doesn't match any of the fire year
    #  y <- x.sub %>% group_by(ptid) %>% 
     #   filter(!year(modis_date) %in% fire_year)
    #  y <-  data.frame(y)
  
  # nice, this works. BUT, still doesn't identify the fires that have >1 MODIS reading
  # What do we do with those??
  # get all rows for which the ptid has a modis year that doesn't match the fire year
  #  x.ys <- subset(x.sub, ptid %in% y$ptid)
  
  # get the inverse of this subset: all well-behaved rows
  #  x.normal <- setdiff(x.sub, x.ys)
  

```
 

```{r}

 
  # Get unique fire years and modis years for each ptid
#  tt <- x.sub %>% group_by(ptid) %>% filter(year(modis_date) )
  
  # get rows for which the fire year == modis year
#  test <-  x.sub[which(x.sub$fire_year == year(x.sub$modis_date)), ]
  
  # get the inverse: ptids for which fire year != modis year
#  missing.fires <- setdiff(levels(x$ptid), test$ptid)
  
```




```{r, echo = F, warning=F}

# #####################################################
#  Assemble the df of fires
# #####################################################

  # all we really need is  a row for each unique fire_year and modis_year
  # PLUS...FRP values for the MODIS years
  #     ...previous years get a holder value
  # ptis, lat, long,, evt...all the stuff that doesn't change for each point.
  
 #  Run this for all years
  # Years pre-2002 do not get an frp
  # we only have 1 frp per year, so we can take the corresponding frp for the modis years
  
  # for each ptid..
  # construct a new dataframe from the consistent info...  # plus evt classes.  All the standard stuff.
  # replicate for the number of years 
  # get the unique years for fire years and modis years



  # set up a df to hold all entries
  df.all <- setNames(data.frame(matrix(ncol = 6, nrow = 0)),  c('ptid', 'latitude', 'longitude', 'fire_year', 'frp', 'FIREID'))


  #remove those pesky classes for this exercise
  x.sub <- x[, -which(names(x) %in% names(x[grepl("class", names(x))]))]
  
  # df with only classes and ptid
  x.classes <- x[, which(names(x) %in% c('ptid', names(x[grepl("class", names(x))])))]
  x.classes <- x.classes[!duplicated(x.classes$ptid), ]
  
  # remove all rows in which the modis date preceeds the fire date
  x.sub <- x.sub[which(x.sub$modis_date > x.sub$fire_date), ]

#x.1 <- x.sub[501:1500, ]
#x.1 <- x[x$ptid == 1305,]
#x.1 <- x.sub[order(x.sub$ptid), ]
#x.1 <- subset(x.1, ptid %in% c(927:950))

x.1 <- x.sub
  
recast_fires = function(f) {

  # trim down the factor levels
  f$ptid <- factor(f$ptid)
#  f <- ungroup(f)

  # list of all unique fire and MODIS years. 
  years <- unique(c(f$fire_year, f$modis_year))
  
  #set up a new df with the info that doesn't change
  f.new <- f[1, c("ptid", "latitude", "longitude")] %>% slice(rep(1:n(), each = length(years)))

  # now add the unique info
    # fire years
  f.new <- cbind(f.new, fire_year = years)

  # get the frp corresponding to the MODIS year
  # years preceeding modis era get NA
  # unique is to catch duplicates of modis year/frp, since the same frp is registered for multiple fire years that
  # preceed the modis era
    #test <- merge(f.new, unique(f[ , c('modis_year', 'frp')]), by.x = "fire_year", by.y = "modis_year", all.x = T)  
    
  merge1 <- left_join(f.new, distinct(dplyr::select(f, modis_year, frp)), by = c("fire_year" = "modis_year"))
     
  # join to get fire_id, etc.
  merge2 <- merge1 %>% left_join(distinct(dplyr::select(f, fireid, fire_year), fire_year, .keep_all = T), by = c("fire_year"))
  
  # consolidate wth holder df
   df.all <- rbind(df.all, merge2)
   
   return(df.all) 
  }
  

#  tidyverse application of fcn to subset
lst.all <- x.1 %>% group_by(ptid) %>%
  do(recast_fires(.))


#convert from df of lists to just a df
#df.all <- do.call(rbind.data.frame, lst.all)
df.all <- data.frame(lst.all)
df.all <- df.all[order(df.all$ptid, df.all$fire_year), ]

# ordering by ptid is problematic b/c of factors
df.all$index <- as.numeric(row.names(df.all))
df.all <- df.all[order(df.all$index), ]
#df.keeper <- df.df

```

```{r}
# #####################################################
#  Get fire order for all points
# #####################################################

temp <- get_fire_order(df.all)
temp$ptid <- as.factor(temp$ptid)
```

```{r, echo=FALSE}
# #####################################################
#  Merge with classes
# #####################################################
# now rejoin the classes to the sorted burn info file

x.classes$ptid <- as.numeric(x.classes$ptid)
x.merged <- merge(temp, x.classes, by = c("ptid"), all.x = T)

x <- x.merged

#x.join <- inner_join(temp, x.classes, by = c("ptid"))
```












```{r}
# Plot the points
df.all %>% 
ggplot(aes(longitude, latitude)) + geom_point(color = 'red', size = 1)
```


```{r, echo=F}
  ################
  
  # Remove all within-season burns: entries tagged as reburns but with year_int = 0.
  #x.1season <- subset(x, reburn == 1 & year_int == 0)
  #x <- x[!(x$FID %in% x.1season$FID), ]
  
  # Redo the rank calculate to account for changes
  #x <- get_fire_order(x)
  
  # Only keep burns for which the MODIS date is after the burn date;
  # no sense in getting modis data for burns that haven't happened yet.
  #x <- x[which(x$modis_date > x$fire_date), ]
  # This was moved earlier
  
  # Record the EVT class that had the greatest area in each 1-km MODIS pixel
  x$max_evt_cat <- apply(x, 1, function(y) names(y[grepl("class", names(y))])[which.max(y[grepl("class", names(y))])[1]])
  
  # ...record the number of pixels that class represented
  x$max_evt_area <- apply(x[grepl("class", names(x))], 1, max, na.rm = TRUE)
  
  # ...calculate the proportion of pixels that class represented
  #x$max_evt_prop <- x$max_evt_pix/max_landsat_pixels * 100

  # Remove rows in which max_class == NA; this means none of the classes had a majority
  x <- x[!is.na(x$max_evt_cat), ]

  # Convert to factor
  x$max_evt_cat <- as.factor(x$max_evt_cat)

  # Merge with file of EVT classes to associate a name with each class
  x <- merge(x, evt_classes, by.x = "max_evt_cat", by.y = "evt_number")
  
  # Keep original set of data prior to excluding duplications
  x.original <- x
  
  # Get rid of all individual classes. Keep as separate file s.t. proportions
  # can be analyzed if desired
  x <- x[, -which(names(x) %in% names(x[grepl("class", names(x))]))]

  # Retain only the highest number of burns for a given point,
  # otherwise a given FRP value will be associated with multiple burn states.
  # First ensure file is ordered by ptid, then by burn_num (highest to lowest)
 # x <- x[order(x$ptid, -x$burn_num), ]
  
  # Remove duplicate points; i.e., all but the first-occurring point, which has the highest burn#
 # x <- x[!duplicated(x$ptid), ]

  # Clean up
#  rm(x.1season)
  

# ----------------------------------
# Calculate STATS ----
# ----------------------------------

# Convert to factor
x$burn_num <- as.factor(x$burn_num)

# FRP by burn count (data)
x.frp.summary <- x %>%
                 group_by(burn_num) %>%
                 summarize(mean = mean(frp, na.rm = T),
                           med = median(frp, na.rm = T),
                           n = length(burn_num))

# > x.frp.summary
# # A tibble: 4 x 4
# `burn_num`   mean   med     n
# <fct>                 <dbl> <dbl> <int>
#   1 1                    197.  77.8 59532
# 2 2                      176.  81.4  9807
# 3 3                      176.  75.7   967
# 4 4                      132.  52.2    67

# # this works reliably...just less fun
 # x.frp2 <- ddply(x, .(burn_num), summarize,
 #                mean = mean(MaxFRP),
 #                med = median(MaxFRP))
 
# Round decimals
# means$MaxFRP <- round(means[,2], 1)

# Get # of EVT classes with majorities
x.maxevt <- summary(x$evt_group)

# Get FRP by veg class. 
# x.frp.class <- ddply(x, .(max_class, burn_num), summarize,
#                mean = mean(MaxFRP),
#                med = median(MaxFRP),
#                n = length(burn_num))

x.maxevt.gp <- x %>% 
                  group_by(evt_group, burn_num) %>%
                  summarize (mean = mean(frp, na.rm = T), 
                            med = median(frp, na.rm = T), 
                            n = length(burn_num))

x.maxevt.gp <- data.frame(x.maxevt.gp)
# > x.maxevt.gp
#      evt_group burn_num      mean    med     n
# 1  Agriculture        1  61.61667  41.55    24
# 2  Agriculture        2  68.20000  68.20     1
# 3        Aspen        1  72.00000  22.20     7
# 4       Barren        1 156.94000  98.70    35
# 5       Barren        2  56.60000  56.60     2
# 6   BirchAspen        1 175.08680  74.20  5947
# 7   BirchAspen        2 174.45053  73.10  1508
# 8   BirchAspen        3 199.33527  70.10   241
# 9   BirchAspen        4 205.76897  30.00    29
# 10          BS        1 232.20899  90.70 10338
# 11          BS        2 209.34827  99.50  1243
# 12          BS        3 113.83810  39.90    21
# 13          BS        4  26.80000  26.80     1
# 14      Burned        1  91.93478  47.90    23
# 15      Burned        2  70.94643  39.75    28
# 16      Burned        3  37.79231  17.90    13
# 17  Floodplain        1 144.17485  65.00  2743
# 18  Floodplain        2 150.51377  79.40   530
# 19  Floodplain        3 133.24286  65.20    21
# 20   Grassland        1 238.92504  92.40   611
# 21   Grassland        2 123.14286  68.40    49
# 22   Grassland        3 355.07500 528.40     8
# 23     Hemlock        1 394.35000 394.35     2
# 24       Marsh        1  26.90000  26.90     1
# 25    Peatland        1 112.45367  50.20   667
# 26    Peatland        2 114.57907  57.20    43
# 27   Shrubland        1 196.21391  73.05  4946
# 28   Shrubland        2 139.85748  71.00   936
# 29   Shrubland        3 139.45874  90.50   143
# 30   Shrubland        4  77.36000  79.40     5
# 31        Snow        1  56.46667  48.30     9
# 32        Snow        2 347.00000 171.20    10
# 33      Spruce        1 183.35000 142.40     6
# 34       Swamp        1 164.35455  81.50    33
# 35       Swamp        2 132.98000  70.40    15
# 36      Tundra        1 141.20876  64.10  6327
# 37      Tundra        2 144.37394  74.70  1869
# 38      Tundra        3 155.08300  58.40   253
# 39      Tundra        4  37.86000  24.40    10
# 40       Water        1 136.94788  84.70   165
# 41       Water        2  83.43571  58.35    28
# 42     Wetland        1 182.97015  75.60 14421
# 43     Wetland        2 187.67659  75.40  1803
# 44     Wetland        3 146.00741  73.90   135
# 45     Wetland        4  97.06190  78.30    21
# 46          WS        1 237.44518  85.20 13227
# 47          WS        2 207.15316  96.30  1742
# 48          WS        3 265.42879 126.40   132
# 49          WS        4  44.60000  44.60     1


# Keep only those levels that have at least 10 points
x.maxevt.gp.top <- x.maxevt.gp[x.maxevt.gp$n > 10, ]

# Keep only those classes that have at least 2 burn levels
x.maxevt.gp.top <- x.maxevt.gp.top[!(as.numeric(x.maxevt.gp.top$evt_group) %in% which(table(x.maxevt.gp.top$evt_group) < 2)), ]

# --------------------------------- 
# Stats
# ---------------------------------

# Calculate 2-way ANOVA; evt group and burn num
x.anova <- aov(formula = MaxFRP ~ evt_group * burn_num, data = x)
summary(x.anova)
TukeyHSD(x.anova, which = "burn_num")
TukeyHSD(x.anova, which = "evt_group")


# Calculate anova: FRP vs year_int, evt, and burn_num
x.anova <- aov(formula = MaxFRP ~ year_int * as.factor(evt_group) * burn_num, data = subset(x, reburn > 0))
summary(x.anova)


# Save data and environment settings   
print(paste0("R data file saved to ", file.path(path.r, rdata)))
save.image(file = file.path(path.r, rdata))

# > x.maxevt.gp.top
# evt_group burn_num      mean    med     n
# 6  BirchAspen        1 275.93009 136.30  6929
# 7  BirchAspen        2 320.58424 187.10  1700
# 8  BirchAspen        3 454.13169 216.60   243
# 9  BirchAspen        4 545.50000 298.90    30
# 10  BlkSpruce        1 406.28216 185.80 11627
# 11  BlkSpruce        2 431.71630 228.90  1350
# 12  BlkSpruce        3 228.29524 131.50    21
# 13     Burned        1 120.75217  81.40    23
# 14     Burned        2 113.38966  62.10    29
# 15     Burned        3  84.79167  81.30    12
# 16 Floodplain        1 210.99758 103.20  3351
# 17 Floodplain        2 259.09631 129.10   731
# 18 Floodplain        3 217.75200  65.20    25
# 19  Grassland        1 416.55549 210.35   638
# 20  Grassland        2 270.20000 188.40    49
# 24   Peatland        1 160.21297  92.95  1072
# 25   Peatland        2 203.98716  71.10   109
# 26  Shrubland        1 261.68958 118.60  6238
# 27  Shrubland        2 204.79135 132.20  1271
# 28  Shrubland        3 207.54667 153.25   150
# 33      Swamp        1 141.90727  78.80    55
# 34      Swamp        2 136.20000  70.40    15
# 35     Tundra        1 214.86234 110.90  7275
# 36     Tundra        2 239.91687 128.40  2033
# 37     Tundra        3 227.70849 110.20   271
# 38     Tundra        4  53.79167  25.60    12
# 39      Water        1 172.32896 109.30   183
# 40      Water        2 240.95172 129.00    29
# 41    Wetland        1 291.45410 136.70 18006
# 42    Wetland        2 302.66308 158.50  2137
# 43    Wetland        3 309.50000 216.25   152
# 44    Wetland        4 381.13043 366.10    23
# 45  WhtSpruce        1 392.02319 185.80 14668
# 46  WhtSpruce        2 392.35054 201.30  1850
# 47  WhtSpruce        3 616.58309 389.80   136
# 

